
---
title: Analysis of predator forging behaviour using multivariate linear mixed effects models
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
    highlight: zenburn
    theme: flatly
    df_print: paged
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.align = "center",
  warning = FALSE, 
  message = FALSE
)
```

<br>



# Overview



## Case study

Explain my study briefly

<br>



## The data

Explain the dataset

<br>



## Load the libraries

```{r}

# libraries
library(data.table)
library(brms)
library(ggplot2)

```

<br>



## Load the data

The data is hosted on a public repository on GitHub. We will export it from there.

```{r}

# Type the repository's URL
github <- "https://raw.githubusercontent.com/quantitative-ecologist"
repository <- "predator-foraging-mode-videogames"
folder <- "main/data"

# Import the data in our session
data <- fread(
  file.path(
    github, repository,
    folder, "FraserFrancoetal2022-data.csv"
  ),
  select = c(
    "player_id", "avatar_id",
    "hunting_success", "game_duration",
    "speed", "space_covered_rate", "hook_start_time",
    "prey_avg_speed", "prey_avg_space_covered_rate"
  )
)

# Rename variables
setnames(data, "hook_start_time", "latency_1st_capture")
setnames(data, "speed", "pred_speed")
setnames(data, "player_id", "predator_id")

```

### Inspect the data structure

```{r}
# Print the data
data

# Inspect the number of avatars
length(unique(data$avatar_id))

# Inspect the number of individual predators
length(unique(data$predator_id))

```

### Visualize the distribution of our data
```{r}

# Reshape the data to plot multiple histograms -----------------------------------

# Select the needed variables
dat_hist <- data[
  , .(
    pred_speed, space_covered_rate,
    prey_avg_speed, prey_avg_space_covered_rate,
    latency_1st_capture, game_duration
  )
]

# Reshape the table
dat_hist <- melt(
  data = dat_hist,
  variable.name = "variable",
  value.name = "value"
)



# Plot the histograms ------------------------------------------------------------

p <- ggplot(data = dat_hist, aes(x = value)) +
  geom_histogram(col = "black", fill = "gray") +
  xlab("\nValue") +
  ylab("Frequency\n") +
  facet_wrap(~ variable, scales = "free") +
  theme_bw() +
  theme(panel.grid = element_blank())

# Show figure
p

```

<br>



# Fitting a multivariate linear mixed effects model: Bayesian inference with the package `brms`



## Standardizing our variables

```{r}

#standardize <- function (x) {
#  (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
#}
#
#data[
#  , c("Zgame_duration",
#      "Zspeed",
#      "Zspace_covered_rate",
#      "Zhook_start_time") :=
#  lapply(.SD, standardize),
#  .SDcols = c(5:8)
#]

```

<br>



## Building the model

### Model 1 - Hunting tactics without controlling for prey behaviour

```{r}

# Model for the predator's speed
speed_form <- bf(
  Zspeed ~
  1 +
  (1 | a | character_name) +
  (1 | b | player_id)
) + gaussian()

# Model for the rate of space covered by the predator
space_form <- bf(
  Zspace_covered_rate ~
  1 +
  (1 | a | character_name) +
  (1 | b | player_id)
) + gaussian()

# Model for the latency before the 1st capture
hook_form <- bf(
  Zhook_start_time ~
  1 +
  Zgame_duration +
  (1 | a | character_name) +
  (1 | b | player_id)
) + gaussian()

```


### Model 2 - Hunting tactics controlling for prey behaviour

<br>



## Setting up priors

```{r}

priors <- c(
  # priors on fixed effects
  set_prior("normal(0, 2)",
            class = "b",
            coef = "Zgame_duration",
            resp = "Zhookstarttime"),
  # priors on var. parameters (brms automatically detects half-normal)
  set_prior("normal(0, 1)",
            class = "sd", # applies to all variance parameters
            resp = c("Zspeed", "Zspacecoveredrate", "Zhookstarttime")),
  # priors on the variance-covariance matrices
  set_prior("lkj(2)",
            class = "cor",
            group = "character_name"),
  set_prior("lkj(2)",
            class = "cor",
            group = "player_id")
)

```


## Run the models

```{r}


#mv_model <- brm(speed_form +
#                space_form +
#                guard_form +   
#                hook_form +
#                set_rescor(TRUE),
#                warmup = 500, 
#                iter = 2500,
#                thin = 8,
#                chains = 4, 
#                inits = "0",
#                threads = threading(10),
#                backend = "cmdstanr",
#                seed = 123,
#                prior = priors,
#                control = list(adapt_delta = 0.95),
#               # save_pars = save_pars(all = TRUE),
#                sample_prior = TRUE,
#                data = data)

```