---
title: Sentinelle Nord Advanced Field School in Computational Ecology
subtitle: Predicting optimal foraging behaviour in TrophIE using a nested patch model
author:
  - name: Maxime Fraser Franco
    affiliation: Département des Sciences Biologiques & Centre de la Science de la Biodiversité du Québec, Université du Québec à Montréal
    email: fraser_franco.maxime@courrier.uqam.ca
  - name: Daniel-Fortin
    affiliation: Département de Biologie & Centre d'étude de la forêt, Université Laval
    email: Daniel.Fortin@bio.ulaval.ca
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    css: style.css
    number_sections: yes
    toc: yes
    toc_float: yes
    highlight: zenburn
    theme: flatly
    df_print: paged
    code_folding: show
#bibliography: references2.bib
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.align = "center",
  warning = FALSE,
  message = FALSE
)
```

---

<br>



# Overview

This workshop is part of the [Sentinelle Nord advanced field school in computational ecology](http://dev.sentinellenord.ulaval.ca/en/ecology2023) held in May 2023 in Val-Morin, Québec, Canada.

<br>



# Case study

Explain case study...

<br>



# The within-patch model

Explain the within patch model...

## Model parameters

courte explication ici...

```{r}
Tf <- 30              # game duration
N_in_t0 <- 50         # total number of marbles at a given station
N1_in_t0 <- 10        # marbles of type 1
N2_in_t0 <- N1_in_t0  # marbles of type2

Nsubstrat <- N_in_t0 - (N1_in_t0 + N2_in_t0)

a_in <- (60 / 10)  # searching rate when not handling (marbles that can be drawn / min)
h1_in <- (15 / 60) # handling time prey type 1 (min / marble)
h2_in <- (15 / 60) # handling time prey type 1 (min/marble)

e1_in <- 40 # points gained by consuming a prey of type 1 (points/marble)
e2_in <- 5  # points gained by consuming a prey of type 1 (points/marble)
p1_in <- 1  # if marble of type 1 are taken, p = 1; 0 otherwise
p2_in <- 1  # if marble of type 2 are taken, p = 1; 0 otherwise
```

## Building the model

courte explication ici...

```{r}
E_in_t0 <- ((a_in * (N1_in_t0 / N_in_t0) * e1_in * p1_in) +
            (a_in * (N2_in_t0 / N_in_t0) * e2_in * p2_in)) /
            (1 + (a_in * (N1_in_t0 / N_in_t0) * h1_in * p1_in) +
            (a_in * (N2_in_t0 / N_in_t0) * h2_in * p2_in))


cons1_in_T0 <- (a_in * (N1_in_t0 / N_in_t0) * p1_in) /
               (1 + (a_in * (N1_in_t0 / N_in_t0) * h1_in * p1_in) +
               (a_in * (N2_in_t0 / N_in_t0) * h2_in * p2_in))


cons2_in_T0 <- (a_in * (N2_in_t0 / N_in_t0) * p2_in) /
               (1 + (a_in * (N1_in_t0 / N_in_t0) * h1_in * p1_in) +
               (a_in * (N2_in_t0 / N_in_t0) * h2_in * p2_in))
```

courte explication ici...

```{r}
E_in_t <- E_in_t0
#cumE_in_t <- E_in_t0
N1_in_t <- N1_in_t0
N2_in_t <- N2_in_t0
N_in_t <- N_in_t0
vN1_in_t <- numeric()
vN2_in_t <- numeric()
vN_in_t <- numeric()
vE_in_t <- numeric()
vnum <- numeric()
vdenom <- numeric()
```

<br>

## Run the model

Courte explication ici... (je ne suis pas sûr si c'est ce code qui roule le modèle)

```{r}

for (i in 1:(Tf)) {

  N1_in_t <- N1_in_t -
            ((a_in * (N1_in_t / N_in_t) * p1_in) /
             (1 + (a_in * (N1_in_t / N_in_t) * h1_in * p1_in) +
             (a_in * (N2_in_t / N_in_t) * h2_in * p2_in))
            )

  N2_in_t <- N2_in_t -
             ((a_in * (N2_in_t / N_in_t) * p2_in) /
              (1 + (a_in * (N1_in_t / N_in_t) * h1_in * p1_in) +
              (a_in * (N2_in_t / N_in_t) * h2_in * p2_in))
             )

  E_in_t <- ((a_in * (N1_in_t / N_in_t) * e1_in * p1_in) +
             (a_in * (N2_in_t / N_in_t) * e2_in * p2_in)
            ) /
            (1 + (a_in * (N1_in_t / N_in_t) * h1_in * p1_in) +
             (a_in * (N2_in_t / N_in_t) * h2_in * p2_in)
            )

  N_in_t <- N1_in_t + N2_in_t + Nsubstrat

  num  <- (a_in * (N1_in_t / N_in_t) * e1_in * p1_in) +
          (a_in * (N2_in_t / N_in_t) * e2_in * p2_in)

  denom <- 1 +
           (a_in * (N1_in_t / N_in_t) * h1_in * p1_in) +
           (a_in * (N2_in_t / N_in_t) * h2_in * p2_in)

  vN1_in_t[i] <- (N1_in_t)
  vN2_in_t[i] <- (N2_in_t)
  vN_in_t[i] <- (N_in_t)
  vE_in_t[i] <- (E_in_t)

  vnum[i] <- (num)
  vdenom[i] <- (denom)

}
```

## Visualize the predictions

courte description?...

```{r}

# Extract parameters
E_in <- vnum / vdenom             # rate of point collection
E_in1 <- append((E_in_t0), E_in)

totgain_temp <- cumsum(E_in1)     # gain function (for each min)
totgain <- totgain_temp[1:Tf]     # gain function only during Tf  (to easily correct for poor coding!)
Residencytime <- 1:Tf

# Plot
library(ggplot2)
dat <- data.frame(x = Residencytime, y = totgain)

ggplot(data = dat, aes(x = x, y = y)) +
    geom_line(linewidth = 1.5, color = "dodgerblue") +
    xlab("\nResidency time (units?)") +
    ylab("Total gain (units?)\n") +
    theme_bw(base_size = 15) +
    theme(panel.grid = element_blank())
```

<br>



# The among-patch model

Explain the among-patch model...

## Model parameters

courte explication ici...

200 patches; 30 participants, ~12 ha; 120 000 m2, mean inter-patch distance = 12.25 m (= 1 / (2 * ((200 / 120000)^0.5))) = 315 x 315 m)
**Patch density :** (patches / m2) if N1 = 1/a, N2 = 0.3/a, then type 1 is better than type 2

```{r}
a <- 30 * 23        # area search rate (m2 / min) (i.e. width * speed)
N1 <- 200 / 120000  # density of patch type 1  (patches / m2)
N2 <- N1            # density of patch type 2  (patches / m2)
mus <- 0.45         # survival probability when travelling between patches
muh <- mus          # survival probability when in a patch (i.e. when handling food)
mur <- .9           # survival probability when in refuge
p1 <- 1
p2 <- 1             # p2 = 0 because only one type of resource patch

hx <- 1:Tf          # h: min spent in a patch handing marbles
# encounter (patch/min) * handling a patch (min/patch)
```

expliquer ceci...

```{r}
Trx <- 1:Tf
hpatcha <- c(rep(hx, times = length(Trx)))
totgain1 <- c(rep(totgain, times = length(Trx)))
Tr <- c(rep(Trx, each = length(hx)))
```

expliquer ceci... serait-ce mieux de séparer un peu le code?
```{r}
Alldata1 <- matrix(c(Tr, hpatcha, totgain1), nrow = (length(Tr)), ncol = 3)
colnames(Alldata1) <- c("Tr", "hpatch", "e_intra")

# Expliquer cette section pour voir si je sépare

# Encounter rate (m2 / min * patch / m2 = patch / min) with patch of type 1
# (only this type present)
lambdaAll1 <- a * N1

# Minutes to encounter next patch = 1 / lambdaAll1
Timepatch_enc1 <- 1 / lambdaAll1

# time spent in each patch + time to the next patch
timeto_search_handl1 <- Timepatch_enc1 + hpatcha

# total no of patches visited during Tf
tot_patch_process_during_Tf <- (Tf - Tr) / timeto_search_handl1
Num_patches <- tot_patch_process_during_Tf

# expliquer cette section pour voir si je sépare

# modulo to get min remaining
remaining <- (Tf - Tr) %% timeto_search_handl1
# if +, then can consume the last patch completely
# i.e. 2.7 patches --> 3 patches completely consumed
diff1 <- remaining - hpatcha

# number of patches fully handled
nb_patch_fullyhandl <- ifelse(
    diff1 < 0,
    floor(Num_patches), ceiling(Num_patches)
)

# gain in the last patch is not fully handled
bbbb <- ifelse(diff1 < 0, totgain_temp[remaining], 0)
# search time at the end.
# = 0 if last patch not fully consumed, remaining otherwise
cccc <- ifelse(diff1 < 0, 0, diff1)
# time spent handling last patch if only partially consumed
dddd <- ifelse(diff1 < 0, remaining, 0)

gain_duringTf <- (nb_patch_fullyhandl * totgain1) + bbbb
Ts_inter <- (Timepatch_enc1 * (floor(Num_patches))) + cccc # assume that consumer starts in a patch
Th1 <- (hpatcha * nb_patch_fullyhandl) + dddd

# overall survival probability, given the time spent in different behaviours
# (searching, handling marbles at a station, waiting in a refuge)
mu <- ((Ts_inter * mus + Th1 * muh + Tr * mur) / Tf)

# Expected long term gain rate, given probability of survival
Eall1 <- gain_duringTf * mu

loc1 <- which(Eall1 == max(Eall1))

zz <- matrix(
    c(
     Eall1, gain_duringTf, mu,
     hpatcha, nb_patch_fullyhandl,
     Ts_inter, Th1, Tr
    ),
    ncol = 8
)
colnames(zz) <- c(
    "Eall", "E", "mu", "h",
    "NbPatch_consumed",
    "Ts", "Th", "Tr"
)

# Long-term expected gain, overall risk, handling rate
# number of stations fully consumed
# time spent searching, handling, in refuge
cc1 <- c(
    Eall1[loc1], gain_duringTf[loc1],
    mu[loc1], hpatcha[loc1], nb_patch_fullyhandl[loc1],
    Ts_inter[loc1], Th1[loc1], Tr[loc1]
)
cc1
```


## Visualize the predictions

```{r}

# Extract model parameters
timex <- 1:length(totgain_temp)
timex <- timex + Ts_inter[loc1]
gain <- totgain_temp / timex

dat <- data.frame(x = timex, y = gain)

ggplot(data = dat, aes(x = x, y = y)) +
    geom_line(linewidth = 1.5, color = "firebrick1") +
    xlab("\nTime (units?)") +
    ylab("Gain (units?)\n") +
    theme_bw(base_size = 15) +
    theme(panel.grid = element_blank())
```


# References {.unnumbered}

<div id="refs"></div>